---

output: pdf_document
---

#Project Inferential Statistics - Week 4
###Carlos Monsalve

##Overview:
#####In this project we will investigate the exponential distribution in R and compare it with the Central Limit Theorem. The exponential distribution can be simulated in R with rexp(n, lambda) where lambda is the rate parameter. The mean of exponential distribution is 1/lambda and the standard deviation is also 1/lambda. Set lambda = 0.2 for all of the simulations. We will investigate the distribution of averages of 40 exponentials. Thousand simulations, at least, needed.


```{r}
setwd("~/Data Science/Inferential Statistics")
```

##Package Required:

```{r}
library(ggplot2)
```

##Variables:

```{r}
n<-40
lambda<-0.2
sim<-1:2000
set.seed(900)
mean<-data.frame(x=sapply(sim,function(x) {mean(rexp(n,lambda))}))
head(mean)
```

##Answer Question 1: Sample Mean versus Theoretical.    

###Simulated Mean:

```{r}
simulatedm<-mean(mean$x)
simulatedm
```

###Expected Center:

```{r}
exp<-(1/0.2)
exp
```

###Comparison:

```{r}
simulatedm-exp
```

...Very close after subtracting, which indicates how close they are. 

##Answer Question 2: Sample Variance versus Theoretical Variance  

```{r}
simulateds<-var(mean$x)
simulateds
```


###Expected Center:

```{r}
exp2<-((1/lambda)/sqrt(40))^2
exp2
```

###Comparison:

```{r}
simulateds-exp2
```

...Very close after subtracting, which indicates how close they are.

##Answer Question 3: Distribution  

#####The distribution of the simulation looks normal. The Mean is around 5 and stddev around 0.79. Symmetric, denser in the center and less dense in the tails, 68% of the area of a normal distribution is within one standard deviation of the mean, approximately 95% of the area of a normal distribution is within two standard deviations of the mean.   

```{r}
ggplot(data=mean,aes(x=x))+geom_histogram(aes(y=..density..),fill="green",binwidth = 0.20,color=I('black'))+stat_function(fun=dnorm,args=list(mean=5,sd=sd(mean$x)))
  
```



